How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-02 tse-paper/content.tex
split into separate files per section

% TODO: Introduction structure
% 0) we introduce & define the problem 1) we survey which techniques exist 2) we implement (for a lack of readily available implementations) and compare the most promising ones 3) we draw conclusions for when to use which technique

% TODO: Include term predictive CI, link to FB paper and https://www-sciencedirect-com.tudelft.idm.oclc.org/science/article/pii/S0164121218301730

% TODO: What about the alternate solution, installing plugins to report precise data? Ie., maven plugins
% - Lots of effort
% - not a generic solution
% - how to analyze history where such plugins were not installed
% - unclear which information you want

% TODO It would be futile to compare our techniques with one of the existing manual regex parsers, as (a) as our literature survey showed, their scope is mainly limited to the Java ecosystem and (b) whether they would work or not would be randomly based on whether the project we select exhibits features their regular expressions were trained for. Finally, they are not really automatic

\lstset{
  morekeywords={},
	basicstyle=\ttfamily\scriptsize,
  postbreak=\mbox{\textcolor{blue}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  stringstyle=\color{Plum},
	frame=single,
  extendedchars=false,
  texcl=false,
  aboveskip=\baselineskip,
  belowskip=0pt
}

Continuous Integration (CI) has become a common practice in software
engineering~\cite{hilton2016usage}.  Many software projects use
CI~\cite{hilton2016usage,staahl2014modeling,beller2017oops} to detect
bugs early~\cite{vasilescu2015quality,duvall2007continuous}, improve
developer productivity~\cite{miller2008hundred,hilton2016usage} and
communication~\cite{downs2012ambient}.  CI builds produce logs which
report results of various sub-steps within the build.  These build logs
contain a lot of valuable information for developers and researchers---for
example, descriptions of compile errors, linter warnings or failed
tests~\cite{beller2017oops,seo2014programmers,vassallo2017a-tale}.

CI builds produce logs which report the results of the steps within the build.
The information stored in build logs can and has
already been used for a variety of applications. One prominent dataset
in the space, TravisTorrent, was featured as the MSR Data Challenge
2017~\cite{msr17challenge}. To date, scientist and practitioners alike
have used it and other, proprietary datasets to understand and
cateogrize Continuous Integration build
failures~\cite{islam2017insights}, to do research on testing
practices~\cite{orellana2017differences}, to train classifiers to
predict the build outcome and
duration~\cite{ni2017cost,bisong2017built,machalica2019predictive},
and to investigate static analysis tools in CI
builds~\cite{zampetti2017open}. Therefore, being able to efficiently
and correctly extract information from build logs is paramount to the
future of a variety of fields depending on it.

However, build logs can be verbose and large---sometimes in excess of
50 MB of ASCII text ~\cite{beller2017oops}---making them inadequate
for direct human consumption. Therefore, to support developers and
researchers in efficiently making use of the information within build
logs, we must at least semi-automatically retrieve the chunks of the
log that describe the targeted information.

There are different techniques to retrieve information chunks from CI
build logs. Beller et al. use a rule-based system of regular
expressions to analyze logs from Travis CI~\cite{beller2017oops}.
Such regular expressions are developed by looking at exemplary build
logs.  Vassallo et al.\ wrote a custom parser to gather information
for build repair hints~\cite{vassallo2018un-break}.  Recently, Amar et
al.\ reduced the number of lines for a developer to inspect by
creating a diff between logs from failed and successful
builds~\cite{amar2019mining}.

These approaches have various strengths and weaknesses: Regular
expressions are exact, but tedious and error-prone to
maintain~\cite{michael2019regexes}.  Custom parsers are powerful
though fragile in light of changes in the log structure. Diffing
between failed and successful logs can reduce the information to be
processed, but is at best semi-automatic~\cite{amar2019mining}.

To gain an impression on how researchers are using information from build logs
within their research and how they are gathering chunks from them to support
developers in understanding build outcomes, we conducted a systematic mapping
study.
Our goal was to classify how and what for researchers use build logs as a source
of data (\textbf{RQ1}).

The outcomes of our study shows that there are various attempts to retrieve
specific information from buildlogs. The resulting implementations are rarely
available and therefore rarely reused. Because of the high development overhead
of the custom parsers, many implementations are limited to small number of supported
build tools.

To address this, we implement three promising techniques for build log analysis.
They retrieve specific chunks of targeted information and are configured by
examples provided by the user.
We conduct an empirical study on the \emph{LogChunks} data set~\cite{brandt2020logchunks} to
gauge the performance of these chunk retrieval techniques and characterize
when one of the techniques should be preferred over other alternatives (\textbf{RQ2}).

\begin{simplebox}{Research Questions}
\begin{itemize}
  \item[\textbf{RQ1:}] Which build log analysis techniques exist?
  \item[\textbf{RQ2:}] Which criteria influence the suitability of a chunk
  retrieval technique for build logs?
\end{itemize}
\end{simplebox}

% At the moment there is only anecdotal evidence on the performance of these
% techniques and on when a technique should be preferred over other alternatives.
% Developers and researches currently have little support when choosing which
% technique to use for a task.

% The goal of this article is to investigate different chunk retrieval techniques
% for build logs and describe under which circumstances certain techniques can be
% recommended over others.
% We aim to characterize different chunk retrieval techniques.
% For \textbf{Research Question 1} (\textbf{RQ1}), we analyze which criteria
% influence the suitability of a chunk retrieval technique for CI build logs.

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and targeted log chunks, PBE synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole text for
  the occurrence of specific trigger words.
\end{itemize}

\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* What went wrong:
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Could not determine the dependencies of task `:app:jacocoTestDebugReport'.
> Task with path 'testDebug' not found in project ':app'.
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]

* Try:
  \end{lstlisting}  
  \caption{Example of a log chunk explaining why an Android/Gradle build failed.}
  % TODO Moritz: we can relate this to travistorrent maybe, you told me long ago that this was something your analyzer cannot extract
  \label{lst:chunk-example-1}
\end{figure}

\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

* What went wrong:
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Execution failed for task ':app:connectedAndroidTest'.
> com.android.builder.testing.api.DeviceException: java.lang.RuntimeException: ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]

* Try:
  \end{lstlisting}  
  \caption{Another example of a log chunk explaining why an Android/Gradle build failed.}
  \label{lst:chunk-example-2}
\end{figure}

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition, RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}, escapeinside=//}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
StartExtraction(RegexPair("Colon/{\color{Plum}$\circ$}/Line Separator", "ALL CAPS")) in EndExtraction(RegexPair("/{\color{Plum}$\varepsilon$}/", "Line Separator/{\color{Plum}$\circ$}/Line Separator"))
  \end{lstlisting}  
  \caption{Simiplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example-1,lst:chunk-example-2}}
  \label{lst:prose-program-simplified}
\end{figure}

\lstset{
  language=,
  morekeywords={},
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
--- FAIL: TestSeparator (0.03s)
    repo\_test.go:75: expected entry to be in form of `* [link] - description.`, got '* ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}  
  \caption{Example of a log chunk. \todo{which build tool?}
  \todo{move to where structural category explanation}}
  \label{lst:chunk-example-3}
\end{figure}

The results of our empirical comparison study on the chunk retrieval techniques
show that the structural representation of the targeted chunk greatly influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example-1,lst:chunk-e:xample-2} show two exemplary log chunks
from Android/Gradle build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed. When PBE is trained on these two examples it produces
a regular expression similar to \Cref{lst:prose-program-simplified}. The retrieval
starts after a colon followed by a line separator and before a capital letter.
It ends before two line separators, consistent with the two example chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2}.
This shows, that the characters structuring the text around the chunk are
cruicial for regular expressions to identify the targeted log chunk.
In fact, if the third
example from \Cref{lst:chunk-example-3} with a different structural
representation is added, PBE is not able to syntesize a program and returns
``\texttt{no program found}''.
We express this difference in structural representation through dividing log
chunks into structural categories. The chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2} are in the same structural
category, while the chunk from \Cref{lst:chunk-example-3} is in a different one.

% \textbf{RQ2} asks under which conditions PBE, CTS and KWS are suited to retrieve information from continuous integration build logs.
% \textbf{RQ2} is refined into sub-questions along with the criteria resulting from \textbf{RQ1} and compares their instantiations for the three techniques:
% how many training examples a technique needs to perform best (\textbf{RQ2.1}), how structurally diverse the examples can be (\textbf{RQ2.2}) and how accurate the retrieved output is (\textbf{RQ2.3}).
% To evaluate PBE, CTS and KWS we use the \emph{LogChunks} data set, which encompasses about 800 log files from 80 repositories.
% Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural representation within the log.

Our study of the three techniques on \emph{LogChunks} shows that
\begin{itemize}
  \item PBE yields very accurate results when trained with two examples from a
  single structural category.
  \item CTS shows the best average precision, though precision and recall of a
  retrieval is hard to determine from the given result.
  A small increase in the number of training examples has no noticeable influence.
  Fewer structural categories improve precision and recall of the retrieval.
  \item KWS has the highest recall of all techniques, however much lower precision.
  It is the technique with the best recall when multiple structural categories
  are present in the training examples.
\end{itemize}

% \noindent
% \textbf{Our work contributes:}
% \begin{itemize}
%   \item A tool with three prototypical implementations of chunk retrieval techniques:
%         \begin{itemize}
%           \item program synthesis from examples using the Microsoft PROSE library (PBE),
%           \item a common information retrieval approach using text similarity (CTS), and
%           \item a keyword search approach (KWS).
%         \end{itemize}
%   \item Recommendations for the configuration of each of the investigated chunk retrieval techniques.
%   \item Guidelines on choosing a suitable chunk retrieval technique.
% \end{itemize}

We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in precision and
recall of the chunk retrieval is required.
CTS is well suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by a human.
In cases where the textual representation of the desired information in the log
is unpredictable or varies greatly, KWS seems to be the best choice. However,
its low precision---it extracts a context of multiple lines around a
finding---makes it generally unsuited for automatic on-ward processing and
instead requires a human to further inspect and interpret the output chunk.

\section{Systematic Literature Survey and Mapping}
\label{sec:survey}

\begin{figure}[htb]
        \centering
	\includegraphics[width=\columnwidth, clip]{img/lit_survey.pdf}
	\caption{Our literature selection process, following \cite{petersen2015guidelines}.}
	\label{fig:lit-survey}
\end{figure}

% TODO (Caro): Integrate Kitchenham's guidelines
From our own experience, we know that the field of analyzing build
logs is relatively young, with perhaps the work on TravisTorrent the
most cited and central one. We also know there have been scattered
research efforts on build log analysis prior to it, mostly where the
log analysis was a side-part of the article. To the best of our
knowledge, there has not been an attempt to systematically overview
and survey them, despite their forseeable future and current importance.

\subsection{Study Selection}
% TODO shorten
\Cref{fig:lit-survey} presents an overview of our study selection.
In this first literature overview study on analyzing build logs, we
therefore want to survey as broad a population of scientific material
as possible. Note that we do not limit our sources to specific
Software Engineering venues, as is often done, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's theses,
academic slide decks, and books. We use Google Scholar as opposed to
Scopus or other databases, as it is the search engine for acdemic
articles with the broadest and most current index, including preprints
from sources such as arXiv. One challenge is that no common
nomenclature has been established yet (in fact, this is one of the
goals of this article).

% TODO Caro cite guidelines
Since we cannot take for granted that ``analyzing build logs'' is the
cannonical term that other, isolated works have used, we instead
decdided to leave out the potentially variable ``analysis'' part and
instead focus on variations of the term ``build log''. This has the
advantage of being biunique: We explicitly decided to exclude systems
or event logs (see \Cref{sec:rw}). On the other hand, a Google Scholar
search for ``build log'' returns more than 2.2 million results
(2020-03-15). We therefore refined our search criterions for the four
different spelling possibilities of ``build log'' to exclude obvious
non-related works. This left us with 1,552 combined search results
(2020-03-15). The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

To make handling so many search results feasible, we employed a
two-pass filtering strategy:

First, we filtered out articles based on 1) their title and 2) the
contextual information Google Scholar displayed. We inlcuded only
publications written in a language intelligible to the authors (that
is, English or German), but only disregarded fewer than 1\% of search
hits due to this criterion. Our strategy was to err on the side of
caution, that is rather including a paper than not. We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query. We noticed that some publishers (ACM, Springer,
Google Books) augment a link to an article with a cutomized ID, which
specifies the link source ({\tt casa\_token, oid}). \todo{overly specific?} We automatically
removed it to find all true duplicates. After this process, we were
left with 259 (17\%) of the original search results.

With these 259 remaining articles, we followed a backward snowball
sampling for related work. This added three papers (five before
duplicate elimination) to our literature set, coming to a total of 261
articles. % TODO Caro: update

On the surviving artciles, we did a second filter phase by downloading the
full articles, reading their abstracts, and searching the full text
for the occurrence of ``log''. If these papers showed traces of
working with build logs, we included them in our survey. For the
snowballed set, we included two of the three articles, both of which
stem from the same first author, were borderline-includes, and had
called build logs from the Linux kernel ``build traces.'' In total, we
were left with 90 articles after this search.

\subsection{Article Classification}
% TODO Caro: please write a draft for this :-)
Having trimmed down the set of articles, we investigated them in depth.
kept all articles that use logs from software builds as input to some kind of process.
we wanted to characterize how researchers use buildlogs as input to any processes.
Interesed in 
\begin{itemize}
  \item what kind of build logs they are using and the source of these logs.
  \item what information they are retrieving from the logs, 
  its kind (chunk/classification) and 
  what they use the retrieved information for
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their tools
\end{itemize}
To objecifiy our data extraction we created a template of the targeted factors
and performed a pilot on five articles. Two authors extracted the data from the
five pilot articles and we held a consensus meeting to unify our understanding
of the factors in the data extraction templates. The remaining articles were
divided over these two authors. We inspected the articles as closesly as necessary,
starting from occurences of ``log'' or ``build'' within the text.
The full data extraction template and our results of this step can be found in
our replication package \todo{cite}.

In this pass we further identified x duplicate articles \todo{do we push those to the last pass?}.
For several cases, multiple articles extended others, though still reported on
the same work with the respect to analyzing build logs.
Of these we only considered the earliest article which reported on the processing
of build logs.

\subsection{Results}
lots of:
- categories + frequencies of answers

\subsubsection{Retrieved Information}

\subsubsection{Retrieval Technique}

\subsubsection{Build Logs}

\subsection{Discussion}
various researchers analyzing build logs to support their research,
tools developed to help developers \todo{motivational reference}.

\subsubsection{Retrieved Information}
why from build logs? not avaliable in cleaner format much use for
research, a bit of developer information / tools following: research
so many logs that unfeasible to do manually, less focus on tools for
devs (caveat! we excluded books and documentation on industrial tools
-> many approaches there to make build logs digestible, tough many
focus on general minimizing of the data presented, not retrieving
specific chunks)

\subsubsection{Retrieval Technique}
broadly mentioned shortly, rarely explained in detail we follow that
researchers often see this as an unimportant side fact work needed to
be done to get the big work done

\subsubsection{Build Logs}
often focussed on specific logs, sometimes mentioned because of effort
developing a parser other times also because other tools used in
further research are limiting the scope (reference paper with PIT
tool). many focus on maven because of it's prevalence also in respect
to other tools much research done on travis ci logs b/c of their
abundance and open availability

\section{Related Work}
% TODO reorder rw by build retrieval technique rather than by paper
\label{sec:rw}
We explain how researchers gather information about CI usage in software projects
through build log analysis and why the chunk retrieval techniques we investigate
simplifies their data collection.
This section moves on to past works about augmenting build logs to make it easier
for developers to inspect them.
The presented approaches are related to and can benefit from the techniques we
analyze in this article.
Further, we classify build logs as semi-structured data and differentiate our
work from system log analysis.

\subsection{Build Failures in CI}
Various researchers studied why CI builds fail and into the impact of
build failures on the development workflow.

Seo et al.~\cite{seo2014programmers} find that a small group of error
types such as dependency mismatches are the most prominent cause of
build failures at Google. In addition, they notice that most failures
are resolved within two builds. They based their analyses on sets of
build logs collected from industry partners. They develop a custom
parser to classify error messages reported by Java and C++ builds.

Rausch et al.~\cite{rausch2017empirical} analyze CI builds of open
source Java projects and find that most builds fail because of failing
tests. For most projects, over half of the failed builds follow a
previous failed build. Rausch et al.'s data shows that most failures
occur in the second half of the build runtime, which can cause long
delays in the feedback loop, especially when builds are automatically
retried upon failure.

Vassallo et al.~\cite{vassallo2017a-tale} compare open source projects
in Java to industrial ones. They determine that testing failures are
more common than compilation errors. Open source builds fail most
often because of unit tests, whereas release preparations are the
primary cause in industrial projects. Their data stems from analyzing
build logs from \emph{TravisTorrent}~\cite{beller2017travistorrent}
with regular expressions.

Ghaleb et al.~\cite{ghaleb2019studying} aim to identify noise in build
breakage data. They classify build failures from \emph{TravisTorrent}
according to whether they were caused by an environmental failure or
caused by a developer change. Their analysis starts with manual
categorization of build logs. They select keywords and strings that
identify their targeted categories and code a script to automatically
classify logs based on these keywords.

\noindent
\textbf{Supporting Log Analysis with Chunk Retrieval}
To leverage the valuable information within build logs the researchers
presented in this section create parsers and regular expression-based
programs. This task of retrieving specific chunks of text from the
build logs can be solved by the chunk retrieval techniques we compare
in this article. Our results can support researchers in choosing a
suitable technique for their data set of build logs and the chunks
they want to retrieve. By relieving them from building custom parsers
we enable them to cover a much wider range of languages and build
tools in their studies.

\subsection{Augmentation of Build Logs}
\label{sec:rw-bl-analysis}
Build logs are a valuable data source for developers to find out why
their build failed. Several researchers investigate how to support
developers in working with verbose build logs. Vassallo et
al.~\cite{vassallo2018un-break} try to shorten the time it takes
developers to understand build logs. They parse
Maven~\cite{maven2019website} build logs into a structured
representation and create hint generators.
%The hint generators leverage this structured access to the
%information within the build log to propose fixes.
For example, one of the hint generators queries stack overflow for
discussions related to why the build failed. In a qualitative study
they observed that highlighting the locality and context of an issue
is helpful to programmers.
%Their tool BART is published as a Jenkins
%Plugin~\cite{bart2019plugin}.
The chunk retrieval techniques we compare in this article can be used
to fill similar structured representations with information from build
logs. As they simplify the construction of parsers they would enable
developers and researchers to cover a wider array of build tools,
which is the main influence factor on the structure of a build log.

Amar et al.~\cite{amar2019mining} compare different approaches to
reduce the portions of a log that a developer has to inspect. Their
techniques remove lines that appear both in logs from passing and
failing builds and use a modified \emph{Term Frequency Inverse
Document Frequency} (TF-IDF) weighting to identify term vectors likely
to occur with failures. Their diff technique can be interpreted as a
chunk retrieval technique, where the targeted information is defined
by the past failures used as basis for the weighted term vectors.

Travis itself already provides build log augmentation in the form of a
basic structuring of the build logs within their log viewer using
\emph{log folds}~\cite{travis2019logfolds}. They add fold identifiers
around common commands and setup or teardown build phases and collapse
the contained lines by default.

\subsection{System Log Analysis}
\label{sec:log-analysis}
A related field of log processing is the processing of system log
files produced during runtime. A main difference between build logs
and system logs is that system logs are fundamentally structured
through events. Each line in a log file represents one event with a
set of fields: timestamp, verbosity level and raw message
content~\cite{he2017towards}. Figure \ref{lst:system-log} shows
example lines from a system log.

The first goal in analyzing system log files is ubiquitously to
separate constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}. Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters. The output of a log
parser is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}. This structured log is then
the input to various machine learning and data mining processes.
%Researchers mine patterns for operational
%profiling~\cite{nagappan2009efficiently},
%debugging~\cite{oliner2012advances}, performance analytics or anomaly
%detection~\cite{nagappan2010abstracting}. Xu et
%al.~\cite{xu2009detecting} leverage the connection of log statements
%to the source code producing them to separate messages into constant
%and variable parts more accurately.

The techniques developed for system log analysis can also be applied
to build logs. One example is comparing execution traces to reference
traces of intended behavior to detect anomalies. Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
%Classic log parsers interpret the whole log file into a sequence of
%events. A similar approach could also be applied to build logs to
%determine the sequence of executed build steps or phases.
In this article, we focus on extracting a single specified information
from the build log as a whole with chunk retrieval techniques. Chunk
retrieval techniques are used as a part of log parsing to retrieve the
values of variable parts in a log message, e.g. by using regular
expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
  keywordstyle=\color{Plum},
  escapeinside=**
}
\begin{figure*}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet Responder: Received block blk_3587508140 of size 67108 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140 terminating
  \end{lstlisting}  
  \caption{System Log excerpt. Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

\subsection{Program Synthesis by Example}
\label{sec:rw-prose}
% \section{Information Extraction and Retrieval Techniques} The
% techniques we investigate are based on existing methods of
% Programming by Example.%, information extraction and information
% retrieval. This section presents different Programming by Example
% resources surrounding the PROSE library. We explain its generic
% program synthesis algorithm and how PROSE synthesizes text
% extraction programs, the foundation of PBE\@.

Programming by Example (PBE) is a form of generalized program
synthesis that enables users to automate repetitive tasks, without
them having to explicitly codify a computer program. One has to
provide prototypical examples of the input and the desired output. A
synthesis algorithm tries to create a program (based on a regular
language) that transforms the input to the output. The minimum working
condition is for this when replayed on the examples inputs is to
provide the specified output, but, hopefully, PBE is as close as
possbile to the user's intention and thus also works on a different
and larger set of inputs. This section introduces the theoretical
foundations of the program synthesis algorithm of the \emph{PROgram
Synthesis using Examples} (PROSE) framework~\cite{prose2019webpage}.
The PROSE framework is developed by Microsoft Research. Next, this
section presents the FlashExtract DSL, which defines text extraction
tasks within PROSE and is the basis for the implementation of our
chunk retrieval technique PBE\@.

\subsubsection{FlashMeta: Inductive Program Synthesis}
The FlashMeta framework presented by Polozov and
Gulwani~\cite{polozov2015flashmeta:} is the backbone of the program
synthesis in the Microsoft PROSE framework. FlashMeta separates the
inductive synthesis algorithm from the domain specific capabilities of
the desired program by encoding the possible program space in a domain
specific language (DSL). The user specifies the desired program
behavior by providing in/output examples (I/O examples). FlashMeta
uses \emph{witness functions}, provided by the DSL, to divide the
synthesis into smaller subtasks. For each of these subtasks it
enumerates all possible programs that solve the subtask consistent
with the set of I/O examples. A program is consistent with a set of
I/O examples if, for each input example, it produces the corresponding
output~\cite{mitchell1982generalization}. The possible subprograms are
joined and stored in a \emph{version space algebra}
(VSA)~\cite{mitchell1982generalization}. This a tree structure, which
space-efficiently saves candidate programs for tasks by sharing common
subexpressions. Next, FlashMeta ranks the enumerated programs
according to which ones the user most likely intended. The DSL also
provides the ranking characteristics. From the ranked VSA, FlashMeta
can then return a ranked list of complete programs consistent with the
user's example.

% In addition to I/O examples of the intended program, the user can
% also provide examples with only input or negative input examples.
% Negative input examples should not be processed by the synthesized
% program.

% The different applications of PROSE presented in the following
% paragraphs were all eventually implemented as DSLs for the FlashMeta
% synthesis algorithm.

\subsubsection{FlashExtract: Data Extraction by Example}
Le et al.~\cite{le2014flashextract:} developed FlashExtract as a DSL
for the Microsoft PROSE framework.
% It enables a user to define text extraction programs for text,
% websites and spreadsheets by giving I/O examples.
FlashExtract's instantiation for text synthesizes extraction programs
from semi-structured text based on regular expressions. Users can
extract multiple fields and structure them with hierarchy and
sequence.
% FlashExtract synthesizes programs to extract each of the fields
% leveraging the information about hierarchical containment and
% sequentiality. It eliminates the need for the user to understand the
% entire structure of the processed document and decreases the effort
% of developing a suitable extraction program.

FlashExtract models the extraction of a single substring as a pair of
two cut positions. A position is either specified by an absolute
character index or by a pair of two regular expressions. The first
regular expression matches the substring directly before the
characterized position, the second regular expression matches the
substring directly after. A regular expression in FlashExtract is a
concatenation of tokens, e.g.\ standard character classes or string
literals frequently occurring in the input examples. Figure
\ref{lst:prose-program} shows a text extraction program synthesized by
FlashExtract. This program defines the first position as after a colon
followed by a newline character and before a piece of text with all
capital letters. It defines the second position as before two newline
characters.

Apart from automatic completion in Excel
spreadsheets~\cite{excel2019flashfill}, FlashExtract is the basis for
two other Microsoft product features: Microsoft's system log analysis
tool Azure Monitor lets users define custom log
fields~\cite{azure2019custom}. The ConvertFrom-String function in
PowerShell allows a user to specify an example template to extract
hierarchical data from a text document~\cite{powershell2019convert}.

We apply the text instantiation of FlashExtract to the domain of build
logs with our chunk retrieval technique PBE\@.

\lstset{
  language=caml,
  morekeywords={PosToEndRegion, RegexPosition, RegexPair, StartToPosRegion},
  keywordstyle=\bfseries\color{black},
  escapeinside=//
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
let s = v in let s = PosToEndRegion(s, RegexPosition(s, RegexPair("Colon/$\circ$/Line Separator", "ALL CAPS"), 1)) in StartToPosRegion(s, RegexPosition(s, RegexPair("/$\varepsilon$/", "Line Separator/{\color{blue}$\circ$}/Line Separator"), 1))
  \end{lstlisting}  
  \caption{Text extraction program synthesized by FlashExtract.}
  \label{lst:prose-program}
\end{figure}

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
We use these techniques to extract text chunks from build logs that
represent a specific, targeted information.

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
In this article, we evaluate different techniques to automatically
retrieve pieces of information that appear literally in the build
logs, i.e., techniques that do not aggreagte, combine, or deducted
information. We call such pieces of information from build logs
\emph{chunks}, and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

With the term \textit{configuration,} we abstract over the training
and parametrization that different techniques require in different
forms. A configuration can be explicitly stated or implicitly dervied
by learning through provided examples. It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log. Each chunk
retrieval technqiue has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word). The
granularity might be adjustable by configuration. \emph{Running a
chunk retrieval technique} means to execute a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output. The array consists of substrings of the
build log text.

\Cref{tab:ctr} shows a comparison of the presented techniques.

\begin{table*}[]
\centering
\caption{Overview of the described chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}XlXlXX@{}} 
\toprule
Name                         & Acronym & Identification Technique                                   & Granularity & Configuration & Source            \\ 
\midrule
Program Synthesis by Example & PBE     & Regular expression program                                 & Character   & In/output examples      \\
Common Text Similarity       & CTS     & TF-IDF \& cosine similarity, expected number of lines & Line        & Output examples           \\
Keyword Search               & KWS     & Keywords, expected number of lines                    & Line        & Keywords, context length  \\
Random Line Retrieval        & RLR     & Random sample                                              & Line        & Retrieval length          \\
Diff Approach                & ---     & Line not present in successful log, information retrieval  & Line        & Logs from failing and successful builds      \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table*}

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
% Text Similarity approaches are used to filter unstructured textual
% software
% artifacts~\cite{runeson2007detection,marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
% One common and simple technique is the Vector Space Model. We
% investigate when text similarity is a suitable technique to retrieve
% information chunks from build logs. In the following we will explain
% the concept of how we apply text similarity to information retrieval
% from CI build logs, which we refer to as \emph{CTS}\@.

\subsubsection{Configuration}
To configure chunk retrieval though text similarity we chose to use
the same concept of examples as for PBE\@
% TODO Caro: More details on this. How do you 'apply' VSM? On what?
and apply the Vector Space Model~\cite{schutze2008introduction}. The
lines of the output strings of the training examples define our search
query. The algorithm splits the search query into single lines and
identifies tokens, in our case words. Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words. Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

\subsubsection{Application}
To retrieve the desired information from a build log, we parse the
whole text and process it in the same way as the output of the
training examples. The algorithm calculates the cosine
similarity~\cite{korenius2007principal} to compare each line of the
build log with each line of the search query. After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity. The average number of
lines in the outputs of the training examples determines how many of
the most similar lines are returned as the output of the retrieval
run.

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad-hoc approach they
use is to search for related keywords. Indeed, this was one of the
most common approaches we took when searching for the reason the build
failed within a log while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.
% As this is a technique readily available in many tools developers
% use to view build logs, we study when such a keyword search is
% suitable for retrieving information chunks from CI build logs. In
% the following we will explain how we use simple keyword search to
% retrieve information from CI build logs, which we refer to as
% \emph{KWS}\@.

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@. To better
compare KWS with PBE and CTS, we also configure it through examples.
We associate each example with keywords which appear in the targeted
chunk or close to it. The configuring keywords for KWS are the ones
that appear most often in the keywords of all training examples.
% TODO Caro Please explain above sentence more precisely

\subsubsection{Application}
For a retrieval run, we take a whole build log file as input and
search for all exact occurrences of the keywords. As keywords are
often not directly describing the desired information, but rather
appear close to the desired information, KWS also retrieves the lines
around the found keyword. The number of surrounding lines retrieved is
the average of lines in the output of the training examples.

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
% \emph{Programming by Example} is a technique which synthesizes
% programs according to in- and output examples provided by the user.
% It enables users to create programs without a priori programming
% knowledge~\cite{mayer2015user}. In the context of text extraction
% through regular expressions, Programming by Example relieves the
% developer from having to understand the whole document structure to
% solve a single extraction task~\cite{le2014flashextract:}. In this
% work, we refer to our interpretation of Programming by Example as
% \emph{PBE}\@. We investigate the suitability of PBE to retrieve
% information chunks from build logs. In the following, we explain the
% configuration and application of PBE to chunk retrieval from CI
% build logs.

>>>>

\input{0-introduction.tex}
\input{1-lit-survey.tex}
%\input{2-rw.tex}
\input{3-techniques.tex}
\input{4-study-design.tex}
\input{5-results.tex}
\input{6-discussion.tex}
\input{7-threats.tex}
\input{8-future-work.tex}
